GSE158422: DNA methylation in Lung squamous cell carcinoma

### 2023-12-20 figs/GSE158422.Rmd
added main script

SERIES_MATRIX="GSE158422_series_matrix.txt.gz"
if(!file.exists(SERIES_MATRIX)){
download.file("https://ftp.ncbi.nlm.nih.gov/geo/series/GSE158nnn/GSE158422/matrix/GSE158422_series_matrix.txt.gz",
  destfile=SERIES_MATRIX)
}
gse <- getGEO(filename=SERIES_MATRIX)
baseDir <- "IDAT"
sample_metadata <- pData(phenoData(gse))
targets <- sample_metadata
mybase <- unique(gsub("_Red.idat.gz" ,"",  gsub("_Grn.idat.gz", "" ,list.files(".",pattern = "GSM",recursive = TRUE))))
targets$Basename <- mybase
head(targets)
rgSet <- read.metharray.exp(targets = targets)
rgSet

@@

sex_female <- as.numeric(colMeans(xdat)  > -0.1)

@@

#MDS plot 1

@@

samplesheet<-targets
groups <- factor(samplesheet$case,levels=c("tumor: normal","tumor: tumor"))
sex <- factor(samplesheet$sex,levels=c("0","1"))
dim(Mval)
dim(beta)

GSE158422 <- dm_analysis(samplesheet=samplesheet,
  sex=sex,groups=groups,mx=Mval,name="GSE158422",
  myann=myann ,beta= beta)

@@

saveRDS(GSE158422,file="GSE158422.rds")
write.table(GSE158422$dma,file="GSE158422_limma.tsv",sep="\t",quote=FALSE)

>>>>

SERIES_MATRIX="GSE158422_series_matrix.txt.gz"
if(!file.exists(SERIES_MATRIX)){
download.file("https://ftp.ncbi.nlm.nih.gov/geo/series/GSE158nnn/GSE158422/matrix/GSE158422_series_matrix.txt.gz",
  destfile=SERIES_MATRIX)
}
gse <- getGEO(filename=SERIES_MATRIX)
baseDir <- "IDAT"
sample_metadata <- pData(phenoData(gse))
sample_metadata$patient <- factor(sapply(strsplit(sample_metadata$title,"_"),"[[",1))

targets <- sample_metadata
mybase <- unique(gsub("_Red.idat.gz" ,"",  gsub("_Grn.idat.gz", "" ,list.files(".",pattern = "GSM",recursive = TRUE))))
targets$Basename <- mybase
head(targets)
rgSet <- read.metharray.exp(targets = targets)
rgSet

@@

sex_female <- as.numeric(colMeans(xdat)  > -0.1)
names(sex_female) <- colnames(xdat)

@@

MDS plot 1

@@

samplesheet <- targets

patient <- factor(sapply(strsplit(samplesheet$title,"_"),"[[",1))

tumor <- factor(samplesheet$`tumor:ch1`,levels=c("normal","tumor"))

sex <- sex_female

dim(Mval)
dim(beta)

design <- model.matrix(~ sex + patient + tumor )
mx <- Mval_flt
name="control vs tumor"

mxs <- mx[,which( colnames(mx) %in% samplesheet$Basename )]
fit.reduced <- lmFit(mxs,design)
fit.reduced <- eBayes(fit.reduced)
summary(decideTests(fit.reduced))
dm <- topTable(fit.reduced,coef=ncol(design), number = Inf)
dma <- merge(myann,dm,by=0)
dma <- dma[order(dma$P.Value),]
dm_up <- rownames(subset(dm,adj.P.Val<0.05 & logFC>0))
dm_dn <- rownames(subset(dm,adj.P.Val<0.05 & logFC<0))
sig <- min(length(dm_up),length(dm_dn))

confects <- limma_confects(fit.reduced, coef=3, fdr=0.05)

if (sig > 10) {
  dmr <- run_dmrcate(mx=mxs,design=design)
  head(dmr)
  comp <- compartment_enrichment(dma)
  comp
  cgi <- cgi_enrichment(dma)
  cgi
} else {
  dmr <- NULL
  comp <- NULL
  cgi <- NULL
}
#make_dm_plots(dm = dm ,name=name , mx=beta,mxs=mxs, groups = groups,
  confects=confects,dmr = dmr, comp=comp, cgi=cgi)

GSE158422 <- list("dma"=dma, "dm_up"=dm_up, "dm_dn"=dm_dn, "confects"=confects,
  "dmr"= dmr, "comp"=comp, "cgi"=cgi, "fit"=fit.reduced)

@@

saveRDS(GSE158422,file="GSE158422.rds")
saveRDS(design,file="GSE158422_design.rds")
saveRDS(mx,file="GSE158422_mx.rds")

write.table(GSE158422$dma,file="GSE158422_limma.tsv",sep="\t",quote=FALSE)
